---
title: "Homework 2"
author: "Saryu Patel"
date: "10/4/2021"
output: github_document
---

```{r setup, message=FALSE}
library(tidyverse)
library(readxl)
```

### Problem 1

```{r}
trash_df <- 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", range = "A2:N534") %>% 
  janitor::clean_names() %>% 
  drop_na() %>% 
  mutate(sports_balls = round(sports_balls))
```

```{r}
precip_18 <- 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 7, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2018)

precip_19 <- 
  read_excel("data/Trash-Wheel-Collection-Totals-7-2020-2.xlsx", sheet = 6, range = "A2:B14") %>% 
  janitor::clean_names() %>% 
  mutate(year = 2019)

precip_df <- 
  bind_rows(precip_18, precip_19) %>% 
  mutate(month = month.name[month])
```

#### Mr. Trash Wheel Dataset

There are `r nrow(trash_df)` observations and `r ncol(trash_df)` variables in the Mr. Trash Wheel dataset. Some key variables in this dataset include the weight of the trash in each dumpster in tons and the volume of the trash in cubic yards. The median number of sports balls in a dumpster in 2019 was `r trash_df %>% filter(year == 2019) %>% summarize(median(sports_balls))`.

#### Precipitation Dataset

There are `r nrow(precip_df)` observations and `r ncol(precip_df)` variables in the resulting precipitation dataset. The key variable in this dataset is the total amount of precipitation each month. The total precipitation in 2018 was `r precip_df %>% filter(year == 2018) %>% summarize(sum(total))` inches, while the total precipitation in 2019 from January through June was `r precip_df %>% filter(year == 2019) %>% summarize(sum(total))` inches.


### Problem 2

Politicians Dataset

```{r message=FALSE}
pols_df <- 
  read_csv("data/fivethirtyeight_datasets/pols-month.csv") %>% 
  separate(mon, into = c("year", "month", "day")) %>% 
  mutate(month = month.abb[as.numeric(month)],
         president = recode(prez_dem, `1` = "dem", `0` = "gop")) %>% 
  select(-c(day, prez_gop, prez_dem))
```

S&P Stock Dataset

```{r message=FALSE}
snp_df <- read_csv("data/fivethirtyeight_datasets/snp.csv")

dates <- snp_df$date %>% as.Date(., "%m/%d/%y")
dates <- as.Date(ifelse(dates > Sys.Date(), format(dates, "19%y-%m-%d"), format(dates)))
snp_df$date = dates

snp_df <- snp_df %>% 
  separate(date, into = c("year", "month", "day")) %>% 
  mutate(month = month.abb[as.numeric(month)]) %>% 
  select(-day)
```

Unemployment Dataset

```{r message=FALSE}
unemp_df <- 
  read_csv("data/fivethirtyeight_datasets/unemployment.csv") %>% 
  pivot_longer(
    Jan:Dec,
    names_to = "month",
    values_to = "unemp_rate"
  ) %>% 
  janitor::clean_names()

unemp_df$year <- as.character(unemp_df$year)
```

Merge

```{r}
final_df <- 
  pols_df %>% 
  left_join(snp_df, by = c("year", "month")) %>% 
  left_join(unemp_df, by = c("year", "month"))
```

The `pols-month.csv` dataset contains information about politicians and the number of which are democratic or republican over time. Once cleaned, it has `r nrow(pols_df)` observations and `r ncol(pols_df)` variables. The `snp.csv` dataset contains information about the stock Standard & Poor's closing value for each day of observation. After cleaning, this data set has `r nrow(snp_df)` observations and `r ncol(snp_df)` variables. The `unemployment.csv` dataset contains information about the unemployment rate for each month throughout each year included in the dataset. This dataset has `r nrow(unemp_df)` observations and `r ncol(unemp_df)` variables. 


Once these three datasets are combined, the resulting data set contains `r nrow(final_df)` observations and `r ncol(final_df)` variables. The years included in the combined dataset range from `r min(final_df$year)` to `r max(final_df$year)`. Some key variables include `president`, `close`, and `unemp_rate`. Of the observations in the combined dataset, `r final_df %>% filter(president == "dem") %>% nrow()` of the months had a democratic president, and `r final_df %>% filter(president == "gop") %>% nrow()` had a republican president. The mean closing value in 2008 was `r final_df %>% filter(year == 2008) %>% summarize(round(mean(close), 2))` while the mean unemployment rate was `r final_df %>% filter(year == 2008) %>% summarize(round(mean(unemp_rate), 1))` that year.


### Problem 3

```{r message=FALSE}
baby_names <-
  read_csv("data/Popular_Baby_Names.csv") %>% 
  janitor::clean_names()

baby_names$gender <- tolower(baby_names$gender)
baby_names$ethnicity <- stringr::str_to_title(baby_names$ethnicity)
baby_names$childs_first_name <- stringr::str_to_title(baby_names$childs_first_name)

baby_names <- distinct(baby_names)

baby_names$ethnicity[baby_names$ethnicity == "Asian And Paci"] <- 
  "Asian And Pacific Islander"
baby_names$ethnicity[baby_names$ethnicity == "Black Non Hisp"] <- 
  "Black Non Hispanic"
baby_names$ethnicity[baby_names$ethnicity == "White Non Hisp"] <- 
  "White Non Hispanic"
```

Popularity of "Olivia" over time

```{r}
baby_names %>% 
  filter(childs_first_name == "Olivia") %>% 
  select(ethnicity, year_of_birth, rank) %>% 
  pivot_wider(names_from = year_of_birth,
              values_from = rank) %>% 
  relocate(ethnicity, `2011`, `2012`, `2013`, `2014`, `2015`, `2016`)
```

Most popular male name over time

```{r}
baby_names %>% 
  filter(gender == "male", rank == 1) %>% 
  select(ethnicity, year_of_birth, childs_first_name) %>% 
  pivot_wider(
    names_from = year_of_birth,
    values_from = childs_first_name
  )
```

Finally, for male, white non-hispanic children born in 2016, produce a scatter plot showing the number of children with a name (y axis) against the rank in popularity of that name (x axis).

Scatterplot

```{r}
baby_names %>% 
  filter(gender == "male", ethnicity == "White Non Hispanic", year_of_birth == 2016) %>% 
  ggplot(aes(x = rank, y = count)) +
  geom_point() + 
  labs(
    title = "Count of Baby Names vs Rank of Baby Names"
  )
```

